---
title             : "The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices"
shorttitle        : "The Sound of Teaching Music"

author: 
  - name          : "Atsuko Tominaga"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Quellenstraße 51, 1100 Vienna, Austria"
    email         : "Tominaga_Atsuko@phd.ceu.edu"
    
  - name          : "Günther Knoblich"
    affiliation   : "1"
    
  - name          : "Natalie Sebanz"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Cognitive Science, Central European University"

authornote: |

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "teaching, expertise, skill transmission, artistic expression, music"
wordcount         : "X"

bibliography      : [""]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
noextraspace      : yes

documentclass     : "apa6"
classoption       : "man"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
---
```{r setup, include = FALSE}
if (!require("papaja")) {install.packages("papaja"); require("papaja")}
if (!require("here")) {install.packages("here"); require("here")}

# create image folder
if (!file.exists(here("paper/image"))){
  dir.create(here("paper/image"))
}
```

# 1. Introduction

# 2. Experiment 1

## 2.1. Method
```{r questionnaire-1, include = FALSE}
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

### 2.1.1. Participants
### 2.1.2. Apparatus and stimuli
### 2.1.3. Procedure
### 2.1.4. Design and data analysis

## 2.2. Results (Articulation)
### 2.2.1. Interonset Intervals (IOIs)
### 2.2.2. Key-Overlap Time (KOT)
### 2.2.3. Key Velocity (KV)
### 2.2.4. KV Transition Points

## 2.3. Results (Dynamics)
### 2.3.1. Interonset Intervals (IOIs)
### 2.3.2. Key Velocity (KV)
### 2.3.3. KV Transition Points
### 2.3.4. Key-Overlap Time (KOT)

## 2.4. Discussion

\newpage

# 3. Experiment 2
## 3.1. Method
```{r questionnaire-2, include = FALSE}
#source(here("experiment-2/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

### 3.1.1. Participants
### 3.1.2. Apparatus and stimuli
### 3.1.3. Procedure
### 3.1.4. Design and data analysis

## 3.2. Results (Articulation)
### 3.2.1. Interonset Intervals (IOIs)
### 3.2.2. Key-Overlap Time (KOT)
### 3.2.3. Key Velocity (KV)
### 3.2.4. KV Transition Points

## 3.3. Results (Dynamics)
### 3.3.1. Interonset Intervals (IOIs)
### 3.3.2. Key Velocity (KV)
### 3.3.3. KV Transition Points
### 3.3.4. Key-Overlap Time (KOT)

## 3.4. Discussion

# 4. General Discussion

# Acknowledgements
We thank Dávid Csűrös for his help with data collection. This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013)/ERC (European Research Council) grant agreement no. 609819, SOMICS, and by ERC grant agreement no. 616072, JAXPERTISE.

# Declarations of interest
All authors have no conflict of interest.

\newpage

# References


\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
\raggedbottom
